---
title: "Regression Analysis and Diagnostics"
author: "Hriday Baghar"
output:
  pdf_document: default
  html_notebook: default
---

# Loading Dataset
```{r, echo=FALSE}
data <- read.csv("Cleaned News Popularity.csv")
str(data)
```

# Exploring correlated features
```{r, echo=FALSE}
cormat <- cor(dplyr::select_if(data, is.numeric))

library(reshape2)
melted_cormat <- melt(cormat)

library(ggplot2)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
We should drop highly correlated features so as to not run into multi-collinearity. Here are some selections we will make for highly correlated variables:

- rate positive words, rate negative words, polarity: polarity will be used

# Regression model
## Base model
```{r, echo=FALSE}
#Building regression model with initial chosen variables

base.model <- lm(shares ~ Day.of.Week + Category + n_tokens_title + n_tokens_content +
                   num_hrefs + num_imgs + num_videos + average_token_length +
                   global_subjectivity + global_sentiment_polarity +
                   title_sentiment_polarity, data = data)

summary(base.model)
```
Testing model assumptions
```{r, echo=FALSE}
plot(base.model)
```
We observe the following:

- It is not clear if linearity holds, the range of residuals is very high
- Constant varaince assumption does not appear to be met

# Regression model with log transformed target variable
```{r, echo=FALSE}
#Transforming target variable to log scale and repeating model building and diagnostics

log.model <- lm(log10(shares) ~ Day.of.Week + Category + n_tokens_title + n_tokens_content +
                   num_hrefs + num_imgs + num_videos + average_token_length +
                   global_subjectivity + global_sentiment_polarity +
                   title_sentiment_polarity, data = data)
```

Testing model assumptions for target variable converted to log:
```{r, echo=FALSE}
plot(log.model)
```
The assumptions of linearity and constant variance appear to be fulfilled for this model.

Model results:
```{r, echo=FALSE}
summary(log.model)
```
All coefficients were found to be significant other than number of tokens in title, number of tokens in content, and article polarity.

We convert the results to percentage changes and create confidence intervals for the same.
```{r, echo=FALSE}
options(scipen=999)

get_percent_change <- function(coef){
  return ((10^coef - 1)*100)
}
estimates.and.ci <- data.frame(percent_change = get_percent_change(log.model$coefficients), get_percent_change(confint(log.model)))
names(estimates.and.ci) <- c("percent_change_estimate", "lower.ci", "upper.ci")
round(estimates.and.ci,3)
options(scipen=0)
```
# Results

Notable findings from regression analysis, every conclusion must be interpreted holding all other model varaibles at a constant value:

- Impact of day of week of publication on number of shares:
  
  - Articles published on Friday are more popular than those published on Tuesday (6.77% less shares), Wednesday (7.3% less shares) and Thursday (6.5% less shares)
  
  - Articles published on weekends are more popular than those on any other weekday (34.53% more shares on Sunday and 28.62% more shares on Saturday than that of Friday on average)
  
  - This is interesting and is likely due to engagement being high on days when people are generally not working
  
- It appears that World and Entertainment are the least popular categories on Mashable and Social Media and Tech are the most popular categories on Mashable (reference category in model is Business)
  
  - Social Media articles are shared 35.22% more and Tech articles are shared 13.5% more than Business articles on average
  
- It appears articles with longer words on average are not as widely shared. An increase in average length of words in an article by 1 unit decreases the number of shares by 13.59% on average

- Articles that are subjective in nature tend to be more popular. A 1 unit increase in subjectivity score increases the number of shares by 6.22% on average

- It looks like more hyperlinks, images and videos in an article have a positive effect on number of shares. The table outlines the precentage change in shares for one unit increase in hyperlink/image/video which could be why the percentage changes are smaller

- Features related to polarity, number of words in article and title do not appear to have a meaningful impact on the number of shares
  